---
title: "Personal Project"
author: "Cihan Tektunalı"
date: "28 Ekim 2017"
output: html_document
---


##1 Introduction
+ The data set I found ([Young People Survey data set]()) consists of responses of a survey done in 2013. Participants were asked multiple questions in 139 categories and evaluated their opinions in a scale of 1 to 5. For the rest of 11 categories, they selected from specific option sets of regarding category. 

+ The data consists of 1010 rows and 150 columns. 139 columns include integer values of intervals whereas 11 columns include categorical data (string).

+ There are missing values in some columns, so some comparisons may not be properly done without data cleaning. Rows with NA values should be considered to remove before creating visulizations. 

## 1.1 What type of analysis can be done with this data set?

 + Correlations between different answers can be explored to reach deeper insights about people's behaviors. 
 
 + Predictions can be made based on specific set of columns. For example; X aged people having Y habits prefer to perform Z type of behavior or not.
 
 + Missing values can be grouped in a way to predict behavior patterns. For example; unanswered questions shows a tendency or not.
 

## 1.2 Load packages and data

```{r setup, warning=FALSE}

# Download data file from source
#download.file("https://github.com/MEF-BDA503/pj-tektunalic/responses.csv")

# Install tidyverse if not already installed
if (!("tidyverse" %in% installed.packages())) {
    install.packages("tidyverse", repos = "https://cran.r-project.org")
}
# Load tidyverse package
library(tidyverse)
# Load the data into variable d
d=read.csv("responses.csv")
```

## 1.3 Initial visualizations 


+Below are some sample visualizations from this data set:


```{r warning=FALSE}
#Replace empty values with NA
d$Gender[as.character(d$Gender)==""] <- NA

#Create histogram plot with variable and label parameters
ggplot(d, aes(x=Age, fill=Gender))+geom_histogram(binwidth=1, alpha=.5, position="dodge")+labs(y="Participant Count",x="Age",title="Participant Numbers By Age (Gender)")

```
```{r warning=FALSE}
#Replace empty values with NA
d$Education[as.character(d$Education)==""] <- NA

#Create histogram plot with variable, axis and label parameters
ggplot(d, aes(x=Age, fill=Education))+
  geom_histogram(binwidth=1, alpha=.5, position="dodge")+
  scale_x_continuous(breaks = c(seq(from = 10, to = 32, by = 1)))+
  scale_y_continuous(breaks = c(seq(from = 0, to = 160, by = 20)))+
  labs(y="Participant Count",x="Age",title="Participant Numbers By Age (Education Level)")
```

